### 2021-10-12 kampus merdeka/kurikulum.Rmd
draft

---
title: "P2MS-ITB TEACHING FACTORY-NETWORK"
subtitle: "Pengantar Market Research"
author: "Mohammad Rizka Fadhli<br>ikanx101@gmail.com"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r,include=FALSE}
rm(list=ls())
library(dplyr)
```