### 2021-10-12 Semester I/Pemrograman dalam Sains/Minggu VIII/20921004 Mohammad Rizka Fadhli Polynomial Numpy.Rmd
new

---
title: "MENYELESAIKAN POLINOMIAL DENGAN NUMPY"
subtitle: | 
    | Tugas Mata Kuliah SK5003 Pemrograman dalam Sains
    | Numpy Python 3
author: | 
    | Mohammad Rizka Fadhli (Ikang)
    | 20921004@mahasiswa.itb.ac.id
date: "`r format(Sys.time(), '%d %B %Y')`"
fontsize: 12pt
output: 
  pdf_document:
    number_sections: false
    toc: false
    fig_width: 7
    fig_height: 4
    fig_caption: true
    highlight: zenburn
    dev: png
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{20921004@mahasiswa.itb.ac.id}
- \fancyfoot[LE,RO]{hal \thepage}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/209_ITB/Semester I/Pemrograman dalam Sains/Minggu VIII")
rm(list=ls())
library(dplyr)
library(ggplot2)
```

\newpage
\tableofcontents
\newpage
\listoffigures
\newpage

# TASK 1

## Soal

Develop a Python program to evaluate polynomial function $y = x^4 + 4x^2 +7$. Find an appropriate interval of $x$ for which the function evaluation is done and plot the graph.

## Jawab

\newpage

# TASK 2

## Soal

Develop a Python program to evaluate the polynomial function $y = 3x^5 + 6$. Find an appropriate interval of $x$ for which the function evaluation is done and plot the graph.

## Jawab

\newpage

# TASK 3

## Soal

Develop a Python program to evaluate the polynomial function $y = 2x^6 - 1.5x^5 + 5x^4 -6.5x^3 + 6x^2 -3x + 4.5$. Find an appropriate interval of $x$ for which the function evaluation is done and plot the relevant data.

## Jawab

\newpage

# TASK 4

## Soal

Develop a Python program to solve the polynomial function $y = x^4 + 4x^2 + 7$.

## Jawab

\newpage

# TASK 5

## Soal

Develop a Python program to solve the polynomial function $y = 3x^5 + 6$.

## Jawab

\newpage

# TASK 6

## Soal

Develop a Python program to solve the polynomial function $y = 2x^6 - 1.5x^5 + 5x^4 -6.5x^3 + 6x^2 -3x + 4.5$.

## Jawab

# == End == {.unlisted .unnumbered}