### 2021-10-12 Semester I/Pemrograman dalam Sains/Minggu VIII/pr.Rmd
all gambar

---
title: "TUGAS POLINOMIAL CHAPTER 15"
subtitle: | 
    | Tugas Mata Kuliah SK5003 Pemrograman dalam Sains
    | Numpy Python 3
author: | 
    | Mohammad Rizka Fadhli (Ikang)
    | 20921004@mahasiswa.itb.ac.id
date: "`r format(Sys.time(), '%d %B %Y')`"
fontsize: 12pt
output: 
  pdf_document:
    number_sections: false
    toc: false
    fig_width: 7
    fig_height: 4
    fig_caption: true
    highlight: zenburn
    dev: png
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{20921004@mahasiswa.itb.ac.id}
- \fancyfoot[LE,RO]{hal \thepage}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
library(dplyr)
library(ggplot2)

gambar_fx = function(f,a,b,delta){
  temp = data.frame(
    x = seq(a,b,by = delta)
  ) %>% 
    mutate(y = f(x)) %>% 
    filter(!is.nan(y))
  subtitle = paste0("Pada selang [",a,",",b,"]")
  caption = "Digambar dengan R\nikanx101.com"
  temp %>% 
    ggplot(aes(x,y)) +
    geom_line(color = "steelblue") +
    theme_minimal() +
    labs(title = "Grafik f(x)",
         subtitle = subtitle,
         caption = caption)

}

```

\newpage
\tableofcontents
\newpage
\listoffigures
\newpage

# TASK 1

## Soal

__15.1__ Develop a Python program to evaluate polynomial function $y = x^4 + 4x^2 +7$. Find an appropriate interval of $x$ for which the function evaluation is done and plot the graph.

__15.4__ Develop a Python program to solve the polynomial function $y = x^4 + 4x^2 + 7$.

## Jawab

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.retina = 10,fig.caption = "Gambar 15.1"}
f = function(x){x^4 + 4*x^2 + 7}
a = -5
b = 5
delta = .05
gambar_fx(f,a,b,delta)
```

\newpage

# TASK 2

## Soal

__15.2__ Develop a Python program to evaluate the polynomial function $y = 3x^5 + 6$. Find an appropriate interval of $x$ for which the function evaluation is done and plot the graph.

__15.5__ Develop a Python program to solve the polynomial function $y = 3x^5 + 6$.

## Jawab

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.retina = 10,fig.caption = "Gambar 15.2"}
f = function(x){3*x^5 + 6}
a = -5
b = 5
delta = .05
gambar_fx(f,a,b,delta)
```

\newpage

# TASK 3

## Soal

__15.3__ Develop a Python program to evaluate the polynomial function $y = 2x^6 - 1.5x^5 + 5x^4 -6.5x^3 + 6x^2 -3x + 4.5$. Find an appropriate interval of $x$ for which the function evaluation is done and plot the relevant data.

__15.6__ Develop a Python program to solve the polynomial function $y = 2x^6 - 1.5x^5 + 5x^4 -6.5x^3 + 6x^2 -3x + 4.5$.

## Jawab

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.retina = 10,fig.caption = "Gambar 15.3"}
f = function(x){2*x^6 - 1.5*x^5 + 5*x^4 - 6.5*x^3 + 6*x^2 - 3*x + 4.5}
a = -5
b = 5
delta = .05
gambar_fx(f,a,b,delta)
```

\newpage

# == End == {.unlisted .unnumbered}