### 2021-12-23 Semester I/Jaringan dan Pengolahan Data Paralel/Ujian/Laporan/Mohammad Rizka Fadhli 20921004.Rmd
save dulu

---
title: "UJIAN"
subtitle: |
  | SK5004 
  | JARINGAN DAN PENGOLAHAN DATA PARALEL
author: |
  | Mohammad Rizka Fadhli
  | 20921004
date: "`r format(Sys.time(),'%d %B %Y')`"
fontsize: 12pt
output:
  pdf_document:
    number_section: false
    toc: false
    fig_width: 7
    fig_height: 4
    fig_caption: true
    highlight: breezedark
linestretch: 1.25
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/209_ITB/Semester I/Jaringan dan Pengolahan Data Paralel/Ujian/Laporan")
```