### 2022-01-29 Semester II/Data Mining/catatan kuliah.Rmd
update

---
title: "DATA MINING"
subtitle: | 
    | Catatan Kuliah
author: | 
    | Ikang FADHLI
    | ikanx101.com
date: "`r format(Sys.time(), '%d %B %Y')`"
fontsize: 12pt
output: 
  pdf_document:
    number_sections: true
    toc: false
    fig_width: 7
    fig_height: 4
    fig_caption: true
    highlight: tango
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{20921004@mahasiswa.itb.ac.id}
- \fancyfoot[LE,RO]{hal \thepage}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/209_ITB/Semester II/Data Mining")
```

\newpage
\tableofcontents
\newpage
\listoffigures
\newpage

# CHAPTER II {-}

# _CONCEPTS AND TECHNIQUES_

## _Types of Datasets_

- Record
    - Relational records
    - Data matrix
    - Transaction data
- Graph and network
    - World wide web
    - Spcial or information networks
- Ordered
- Spatial, image and multimedia

## Important Characteristics of Structured Data

- Dimensionality
    - Curse of dimensionality
- Sparsity
    - Only presence counts
- Resolution
    - Patterns depend on the scale
- Distribution
    - Centrality and dispersion