### 2022-05-17 Semester II/Research Method/UAS/tabel only.Rmd
update

---
title: "Untitled"
output: html_document
date: '2022-05-17'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(dplyr)

load("bahan review.rda")
```

```{r}
gabung_all %>% knitr::kable()
```