### 2022-10-05 Semester III/Pengantar SK/PR 2/PR 20921004.Rmd
new

---
title: "PR 2 Pengantar Sains Komputasi"
subtitle: "20921004 Mohammad Rizka Fadhli"
output: pdf_document
date: '5 Oktober 2022'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

# SOAL 1

_Give the magnitude and precision of_ $93.6 \times 10^7$

# SOAL 2

_We do the calculation of_

$$\begin{matrix}
x = 0.536 \\
y = 7 x \\
z = y / 7
\end{matrix}$$

_on the system that truncats to 3 digits of significand_. _Write an IF statement that puts 1 in the value of a variable p if the floating point of x and z are equal and otherwise puts 0_.

# SOAL 3

_Consider a machine that rounds to 4 significant digits. Suppose initially _ $y = 9.649$ _and_ $x = 7.834$. _The following assignment statement, which calculates the expression on the right,_ $(y + x)$ _and the replaces the value of_ $y$ _on the left with the result, is in the loop that executes four times:_ $y = y + x$. _After each iteration of the loop, give the value stored in_ $y$ _and the absolute and relative error._

\newpage

# JAWABAN SOAL 1

\newpage

# JAWABAN SOAL 2

\newpage

# JAWABAN SOAL 3