### 2022-10-05 Semester III/Pengantar SK/PR 2/PR-20921004.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={PR 2 Pengantar Sains Komputasi},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{PR 2 Pengantar Sains Komputasi}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{20921004 Mohammad Rizka Fadhli}
\author{}
\date{\vspace{-2.5em}5 Oktober 2022}

\begin{document}
\maketitle

\hypertarget{soal-1}{%
\section{SOAL 1}\label{soal-1}}

\emph{Give the magnitude and precision of} \(93.6 \times 10^7\)

\hypertarget{soal-2}{%
\section{SOAL 2}\label{soal-2}}

\emph{We do the calculation of}

\[\begin{matrix}
x = 0.536 \\
y = 7 x \\
z = y / 7
\end{matrix}\]

\emph{on the system that truncats to 3 digits of significand}.
\emph{Write an IF statement that puts 1 in the value of a variable p if
the floating point of x and z are equal and otherwise puts 0}.

\hypertarget{soal-3}{%
\section{SOAL 3}\label{soal-3}}

\emph{Consider a machine that rounds to 4 significant digits. Suppose
initially } \(y = 9.649\) \emph{and} \(x = 7.834\). \emph{The following
assignment statement, which calculates the expression on the right,}
\((y + x)\) \emph{and the replaces the value of} \(y\) \emph{on the left
with the result, is in the loop that executes four times:}
\(y = y + x\). \emph{After each iteration of the loop, give the value
stored in} \(y\) \emph{and the absolute and relative error.}

\newpage

\hypertarget{jawaban-soal-1}{%
\section{JAWABAN SOAL 1}\label{jawaban-soal-1}}

\newpage

\hypertarget{jawaban-soal-2}{%
\section{JAWABAN SOAL 2}\label{jawaban-soal-2}}

\newpage

\hypertarget{jawaban-soal-3}{%
\section{JAWABAN SOAL 3}\label{jawaban-soal-3}}

\end{document}