### 2022-11-30 Thesis/Overleaf Version/tex/tex.Rmd

---
title: "Untitled"
author: "Mohammad Rizka Fadhli"
date: "11/29/2022"
output: 
  pdf_document:
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(kableExtra)

rm(list=ls())
load("~/209_ITB/Thesis/Overleaf Version/report.rda")
```

\newpage

```{r,echo=FALSE}
solusi_7 %>% 
  mutate(value = value/1000,
         value = round(value,1)) %>% 
  rename("Raw material" = k,
         "Total order qty (in tons)" = value) %>% 
  select(-variable) %>% 
  janitor::adorn_totals() %>% 
  knitr::kable(caption = "Optimal Solution",align = "c")

df = 
  solusi_8 %>% 
  mutate(value = value/1000,
         value = round(value,1)) %>% 
  reshape2::dcast(k~j,value.var = "value") %>% 
  rename("Raw material" = k)
colnames(df)[2:5] = paste0("Week ",1:4)
df %>% knitr::kable(caption = "Raw Material order per week (in tons)",align = "c")

df = 
  solusi_3 %>% 
  mutate(value = value/1000,
         value = round(value,1)) %>% 
  reshape2::dcast(k~j,value.var = "value") %>% 
  rename("Raw material" = k)
colnames(df)[2:5] = paste0("Week ",1:4)
df %>% knitr::kable(caption = "Stock at the end of the week (in tons)",align = "c")

```