### 2023-03-07 Semester II/Research Method/Manuskrip ke Software X/Manuscript/table only.Rmd

---
title: "Tables"
output: 
  pdf_document:
    keep_tex: true
date: "2023-03-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

```{r,echo=FALSE}
library(readxl)
df = read_excel("~/209_ITB/Semester II/Research Method/Tugas Kuliah I/20201207 Product list.xlsx") %>% janitor::clean_names()
df %>% 
  select(product_code,burn_3m,cost_benefit) %>% 
  rename(budget = burn_3m,
         expected_revenue = cost_benefit) %>% 
  filter(expected_revenue > 0) %>% 
  head(15) %>% 
  knitr::kable(caption = "Sample Portfolio Data")
```

\newpage

```{r,include=FALSE}
load("~/209_ITB/Semester II/Research Method/Paper Finale/bahan review.rda")
```

```{r,echo=FALSE}
gabung_all %>%
  select(contains("budget")) %>% 
  mutate(eksak_budget = round(eksak_budget/5000000 * 100,2),
         existing_budget = round(existing_budget/5000000 * 100,2),
         spiral_budget = round(spiral_budget/5000000 * 100,2)) %>% 
  mutate(dataset_ke = 1:10) %>% 
  relocate(dataset_ke,.before = eksak_budget) %>% 
  rename("#portfolio" = dataset_ke,
         exact_budget = eksak_budget) %>%
  knitr::kable(caption = "Budget Fullfilment: What percentage of the budget is spent on the solution?")
```

\newpage

```{r,echo=FALSE}
gabung_all %>%
  select(contains("revenue")) %>% 
  mutate(existing_revenue = round(existing_revenue/eksak_revenue * 100,2),
         spiral_revenue = round(spiral_revenue/eksak_revenue * 100,2)) %>% 
  mutate(dataset_ke = 1:10) %>% 
  relocate(dataset_ke,.before = eksak_revenue) %>% 
  mutate(eksak_revenue = round(eksak_revenue / 1000000,3),
         eksak_revenue = paste0(eksak_revenue," mil IDR"),
         existing_revenue = paste0(existing_revenue,"%"),
         spiral_revenue = paste0(spiral_revenue,"%")) %>% 
  rename("#portfolio" = dataset_ke) %>% 
  knitr::kable(caption = "Total Expected Revenue: What is the percentage of revenue achieved compared to the revenue of the exact solution?")
```

\newpage

```{r,echo=FALSE}
sama_tdk = 
  data.frame(dataset = 1:10,
             banyak_produk_sama = sama,
             banyak_produk_beda = 100-sama) %>% 
  rename(similar_product = banyak_produk_sama,
         dissimilar_product = banyak_produk_beda) %>% 
  rename("#portfolio" = dataset)
sama_tdk %>% knitr::kable(caption = "Similarity Product Portfolio")
```

\newpage

```{r,echo=FALSE}
data.frame(dataset = 1:10,
           eksak_y = ompr_sum,
           eksak_n = 100 - ompr_sum,
           spiral_y = spiral_sum,
           spiral_n = 100 - spiral_sum
           ) %>% 
  mutate(portofolio_eksak = paste0("Discount: ",eksak_y,
                                   "; No discount: ",eksak_n),
         portofolio_spiral = paste0("Discount: ",spiral_y,
                                   "; No Discount: ",spiral_n)) %>% 
  rename("#portfolio" = dataset) %>% 
  select(-eksak_y,
         -eksak_n,
         -spiral_y,
         -spiral_n) %>% 
  knitr::kable(caption = "Product Proportion: How many products are given and not discounted?")
```