### 2023-04-11 Thesis/MethodX/Manuscript/versi doc.Rmd
update paper

---
title: "Untitled"
bibliography: mybibfile.bib
output: pdf_document
date: "2023-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Results

In this section, we give an example of the solution of the optimization model we derived in the previous section. In this example, we consider an instant where $N=6$ and $I=51,$ so that we have 2508 decision variables.

## Parameters values

The following tables show the matrix of raw-material flexibility of all items.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
rm(list=ls())

library(dplyr)
library(tidyr)
library(readxl)
library(kableExtra)

```
\pagebreak

```{r,echo=FALSE,message=FALSE,warning=FALSE}
rm(list=ls())

library(readxl)
library(kableExtra)

df = read_excel("~/209_ITB/Thesis/Overleaf Version/data input.xlsx",sheet = "matriks gula x produk")
df[is.na(df)] = 0
colnames(df)[2:7] = 1:6
df[1:15,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 1-25") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"Raw material" = 6))
```

```{r,echo=FALSE,message=FALSE,warning=FALSE}
df[16:51,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 26-51") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"Raw material" = 6))
```

Next, we present the instant of the total raw material demands during a planning horizon.

```{r,echo=FALSE,message=FALSE,warning=FALSE}
rm(list=ls())

my_comma = scales::label_comma(accuracy = .1, big.mark = ",", decimal.mark = ".")

df = read_excel("~/209_ITB/Thesis/Overleaf Version/data input.xlsx",sheet = "matriks produk x minggu")
colnames(df)[2:5] = paste0("week ",1:4)

df$`week 1` = my_comma(df$`week 1`)
df$`week 2` = my_comma(df$`week 2`)
df$`week 3` = my_comma(df$`week 3`)
df$`week 4` = my_comma(df$`week 4`)

df[1:10,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 1-11") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"raw -material demand (in kg)" = 4))
```

\newpage

```{r,echo=FALSE,message=FALSE,warning=FALSE}
df[11:51,] %>% 
  rename(item = code_product) %>% 
  knitr::kable(caption = "the flexibility matrix of items 12-51") %>% 
  kable_styling(latex_options = "hold_position") %>% 
  add_header_above(c(" " = 1,"raw-material demand (in kg)" = 4))
```